---
title: "2301-353-353m-453-LE "
subtitle: "Profs: R. H. French, L. S. Bruckman, P. Leu, K. Davis, S. Cirlos"
author: "TAs: W. Oltjen, K. Hernandez, M. Li, M. Li, D. Colvin" 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    latex_engine: xelatex
    toc: true
    number_sections: true
    toc_depth: 6
    highlight: tango
  html_document:
    toc: yes
urlcolor: blue
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  cache = FALSE, # if TRUE knitr will cache results to reuse in future knits
  fig.width = 6, # the width for plots created by code chunk
  fig.height = 4, # the height for plots created by code chunk
  fig.align = 'center', # how to align graphics. 'left', 'right', 'center'
  dpi = 300, 
  dev = 'png', # Makes each fig a png, and avoids plotting every data point
  # eval = FALSE, # if FALSE, then the R code chunks are not evaluated
  # results = 'asis', # knitr passes through results without reformatting
  echo = TRUE, # if FALSE knitr won't display code in chunk above it's results
  message = TRUE, # if FALSE knitr won't display messages generated by code
  strip.white = TRUE, # if FALSE knitr won't remove white spaces at beg or end of code chunk
  warning = TRUE, # if FALSE knitr won't display warning messages in the doc
  error = TRUE) # report errors
  # options(tinytex.verbose = TRUE)
```

License: Roger H. French, Copyright 2023, All Rights Reserved.

\setcounter{section}{2}
\setcounter{subsection}{1}
\setcounter{subsubsection}{0}

#### Lab Exercise 2a: Classifiation

2 Questions, for a total of 2 points

  - Question 1 = 1.5 points
  - Question 2 = 0.5 points

#### Question ISLR Chapter 4, Exercise 10


    
##### (a) 

```{r}

```

Answer: 



#### Question ISLR Chapter 4, Exercise 12



##### (a) Write a function, Power(), 

  
  
```{r}

```

Answer:   



#### Links
 


<!--
# Keep a complete change log history at bottom of file.
# Complete Change Log History
# v0.00.00 - 1405-07 - Nick Wheeler made the blank script
##########

