---
title: "Class and Coding Expectations (CWRU, Pitt, UCF, UTRGV)"
subtitle: "Profs: R. H. French, L. S. Bruckman, P. Leu, K. Davis, S. Cirlos"
author: "TAs: W. Oltjen, K. Hernandez, M. Li, M. Li, D. Colvin" 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    latex_engine: xelatex
    toc: TRUE
    number_sections: TRUE
    toc_depth: 6
    highlight: tango
  html_notebook:
  html_document:
    css: ../lab.css
    highlight: pygments
    theme: cerulean
    toc: yes
    toc_depth: 6
    toc_float: yes
    df_print: paged
urlcolor: blue
always_allow_html: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  cache = FALSE, # if TRUE knitr will cache results to reuse in future knits
  fig.width = 6, # the width for plots created by code chunk
  fig.height = 4, # the height for plots created by code chunk
  fig.align = 'center', # how to align graphics. 'left', 'right', 'center'
  dpi = 300, 
  dev = 'png', # Makes each fig a png, and avoids plotting every data point
  # eval = FALSE, # if FALSE, then the R code chunks are not evaluated
  # results = 'asis', # knitr passes through results without reformatting
  echo = TRUE, # if FALSE knitr won't display code in chunk above it's results
  message = TRUE, # if FALSE knitr won't display messages generated by code
  strip.white = TRUE, # if FALSE knitr won't remove white spaces at beg or end of code chunk
  warning = FALSE, # if FALSE knitr won't display warning messages in the doc
  error = TRUE) # report errors
  # options(tinytex.verbose = TRUE)
```


 \setcounter{section}{4}
 \setcounter{subsection}{2}
 \setcounter{subsubsection}{2}

#### Class Readings, Assignments, Syllabus Topics

##### Reading, Lab Exercises, SemProjects

  - Readings: 
    - For today: R4DS 1-3
    - For next class: OIS5, (EDA 32-58)
  - Laboratory Exercises: 
    - LE2 : Due today at midnight  
    - LE3 : Given out today
  - Office Hours: (Class Canvas Calendar for Zoom Link) 
    - Wednesday @ 4:00 PM to 5:00 PM, Will Oltjen
    - Saturday @ 3:00 PM to 4:00 PM, Kristen Hernandez 
    - **Office Hours are on Zoom, and recorded**
  - Semester Projects
    - DSCI 451 Students Biweekly Update 2 due 9/23
    - DSCI 451 Students 
      - Next **Report Out #1 is Due Friday September 30th**
    - All DSCI 351/351M/451 Students: 
      - **Peer Grading of Report Out #1 is Due October 11th, 2022**
    - Exams
      - MidTerm: Tuesday October 18th, in class or remote, 11:30 - 12:45 PM
      - Final: Monday December 19, 2022, 12:00PM - 3:00PM, Nord 356 or remote

#### Introduction

- Lets go over some of our expectations regarding this course.

This document outlines general expectations moving forth.

Please read through this document carefully and 

  - if you have any questions,
  - do message us on Slack.

##### LE Grading Rubric, as an example

###### Global Points

* -1 pts per day late

###### Question Points

  - 1-1 Completion (½ pt)
  - 1-2 Structure Completion (1 pt)
    - Give the Class of Each Column (1/4)  pts
    - Subset using versicolor (1/4) pts
    - Mean and Median (1/4) pts
    - lm () & plot (1/4)  pts
  - 1-3 OIS Completion (1) pt
    - Cases (1/3) pts
    - Numerical Variables (1/3) pts
    - Categorical (1/3) pts
  - 1-4 Distributions  (1) pt 
    - 1/8 pts answer 
    - 1/8 pts for explanation
    - 1/8 pts answer 
    - 1/8 pts for explanation
    - 1/8 pts answer 
    - 1/8 pts for explanation
    - 1/8 pts answer 
    - 1/8 pts for explanation
  - 1-5 OIS 2.9 (1 pt) 
    - 1/8 pts answer 
    - 1/8 pts for explanation
    - 1/8 pts answer 
    - 1/8 pts for explanation
    - 1/8 pts answer 
    - 1/8 pts for explanation
    - 1/8 pts answer 
    - 1/8 pts for explanation
  - 1-6 For Loops (1/2) pt
    - 1/4 pt Attempt
    - 1/4 pt Success
  - 1-7 Heart Transplants (2) pts
    - 1/3 pts Answer + Reason
    - 1/3 pts Answer + Reason
    - 1/3 pts Answer + Reason
    - 1 pts
      - 1/4  pts Answer
      - 1/4  pts Answer 
      - 1/2  pts Answer + Reason


#### General Expectations

- Please follow these points carefully as you work on your lab exercises.

The grading will be done based on these points.

To provide an overview of the points:-

  - Remember that the knitted PDF/HTML shows the first 80 characters in the code
    - as the output and omits the rest. 
  - Make sure you format your code chunks properly. 
    - Please verify this before submission.
  - Since we'll be grading your work, be sure to comment your code. 
    - Suppose you use a function that is not common (for example, unite()), 
    - explain what it does in a short comment and which package you used. 
  - It would also help us learn about new functions in the process.
  
There are multiple approaches to the same question. 

  - We'll take that into account while grading.
  
Please answer questions outside the code chunk 

  - (next to ANSWER -> or Answer). 
  - Suppose, the question asks you to compare trends, 
    - don't type it as a comment in a code chunk.
  
Be creative while you work on plots. 

  - As you will learn along, ggplot2 package has amazing functionalities. 
    - Spend some time formatting plots to make them more understandable. 
  - For example, 
    - adding a plot title, 
    - axes labels (specify units wherever required), 
    - changing theme specifications will earn full points.
  
Include meaningful variable names. 

  - For a dataset containing billboard lyrics, 
    - billboard_lyrics is too long and bbl is too short. 
  - Find a balance between the two extremes like 'songs'. 

#### Slack

- Please use Slack as your primary mode of communication with us, 

  - especially for asking questions. 
  - This way everyone can learn from the discussion

Before you ask a question, 

  - please verify if the question has already been asked
  - by other students by scrolling up. 

Don't forget to tag us by using @ 

  - because it's highly probable that your messages 
    - might get lost before the submission. 

While asking questions, 

  - refer to the question number while typing your question
  - and show a minimum working example (MWE)
    - (explain what you did and send a screenshot).

If you are facing an error 

  - or have difficulty in debugging or troubleshooting, 
  - please send us a screenshot. 

It'll be really nice to ask questions within threads 

  - and post it on the main channel
  - so that we won't miss them.

Our handles are - 

  - TAs: 
    - @Will Oltjen (CWRU, TA)
    - @Kristen Hernandez (CWRU, TA) 
    - @Mingxuan Li 
  - Professor: 
    - @Roger French
    - @Paul Leu

Try your best to ask your questions in the class channel #dsci351-351m-451. 

  - This way others can also benefit and help with your answers. 
  - It'll be a win-win situation!

If something has to be privately discussed with the TAs or Prof. French, 

  - please DM directly on Slack. 
    -  or contact by email
  - Please do not message LE questions by DMs.

While we will try our best to reply as soon as possible, 

  - please allow for some potential delays in our response rate. 
  - We cannot guarantee answering questions 
    - 10 minutes before the submission deadline, 
    - so make sure you ask questions earlier
    - in order to receive a response from us.

Please be aware that we are a large group of people 

  - and there are few TAs. 
  - Please feel free to form your own study groups 
    - and discuss among yourselves. 
  - The TAs, cannot be your study buddies 
    - and cannot organize personal Zoom sessions 
    - unless it is crucial. 

In the Slack channel, 

  - please don't wait for the TAs to respond to all the questions.
  - If you browse through the Slack channel
    - and you know you can answer a question, 
    - we encourage you to do that. 

We won't be available late at night 

  - just before the submission deadline. 
  - Last-minute questions before the deadline 
    - will probably not be answered until the next morning. 

#### Office Hours

- We have four office hours in the week:- 

  - CWRU: Mondays and Wednesdays at 4:30-5:30 pm 

  - UPitt: Mondays 11 am-12 pm and Tuesdays 1-2 pm

Please attend the office hours if you have questions and any other concerns.

  - These sessions ARE recorded.

Because we have about 10-15 students attending the office hours 

  - and there are fewer TAs, 
  - please be patient and ask questions one by one.

Please make sure to give others the time to ask their questions. 

  - Some students ask several questions at once 
    - without giving others time to ask questions.

#### Submission Time

- Typically lab exercises are expected to be submitted by midnight

  - of the deadline day.

Requests made on the day of submission 

  - will probably not be accommodated.

Make sure you are able to knit your documents earlier before submission. 

If you are facing any challenges or problems 

  - that prevent you from turning in the lab exercise on time, 
    - please contact us with an explanation of why you need extra time 
    - and we'll see if your request can be accommodated.

#### Git Commands

- **NOTE: NOT FOLLOWING THE GIT COMMANDS COULD LEAD TO SEVERAL PROBLEMS.**

I cannot emphasize enough why this is important.

Sameera's personal story: 

  - 'when I was a DSCI student, I wasn't regular at git pushing and git pulling.
  - At one point, I lost all of my files from the ODS Desktop computer
    - and thankfully, it was possible to retrieve 
    - after spending a few hours from the CWRU Network.
  - All of this happened just before the midterm exam!
    - It was definitely scary.
  - Please don't land in such situations!'

Git pushing and pulling regularly 

  - will save your work virtually 
    - and in events of your computer crashing for no particular reason, 
    - you can always clone your fork.
  - And remember **every hard drive or usb stick will fail**. 
    -  You can rely on that

#### OnDemand (Markov, RStudio Server-4.1.1) Versus Your Own Local Computer

- There are several reasons why we encourage you to use ondemand.case.edu.

  - Not just because it is fancy to do virtual computing 
    - but also because your session can be recovered 
    - (given you input a longer duration).

The packages would be updated from time to time 

  - and it's easier to reach out to help@case.edu about the problems.

In fact, we can try diagnosing the issues from our end too.

  - We understand that because of spotty internet connection, 
    - it's can be hard to use virtual computing.
    - But the Rstudio Server's cache keystrokes, which helps
  - In such cases, make sure you have the Git Bash app set up 
    - and do git pushing/pulling regularly.

If you ever face any issue, please contact us.

Additionally, you can send a ticket to:-

  - ODS Desktop problems: help@case.edu
  - Markov problems: hpcsupport@case.edu

#### Coding Style

- Code styling is important!

  - As data scientists, part of our job is 
    - to effectively communicating our process
    - and results to a broader audience. 
  - Martin Fowler: “Any fool can write code that a computer can understand. 
    - Good programmers write code that humans can understand.”
  - Good coding style helps you find bugs and will save you time in the long run.
  - You may collaborate with others on coding.  
    - It is important that others can understand your code.  
  - **YOU MUST BE ACTIVELY MAKING YOUR CODE READABLE**. 
    - As you are coding, use the below practices to make your code easy to understand.  
  - After your code is functional, 
    - do several passes over your code to clean up the code, 
    - simplify it, and ensure it is something another programmer can understand.  
  
We expect your code to be 

  - compact, well-commented and effectively answers 
    - the question you are trying to solve.

There needs to be a whitespace between symbols and characters. 

  - `hits %>% filter(Artist=="madonna")` is not the ideal style
  - `hits %>% dplyr::filter(Artist == "madonna")` is!

There are multiple methods to solve a question 

  - and new approaches to solving a question would always be appreciated.
  
If you are using a new function that is not regularly used, 

  - please use package::function approach 
    - and say what it does. 
  - It can be a comment in a code chunk.
    
Going forward, increasing scrutiny will be placed on the following things

  - (including the general expectations) 
    - and points will be deducted accordingly. 

**NOTE**

When looking at the .Rmd file you can notice something interesting. 

Every sentence I write is ended by a **PERIOD** a **SPACE** and a **ENTER** key.

Although it creates a new line of code, 

  - when it is knitted they will all be in one block of text. 

##### General

* Use RStudio Diagnostics.  Diagnostics can be enabled and set within the 
```Global Options -> Code -> Diagnostics editing pane:```

Do not use magic numbers.  

  - [Magic numbers](https://en.wikipedia.org/wiki/Magic_number_(programming) are numbers that appear directly in computations.  
  - The use of magic numbers 
    - obscures the designer's intent in choosing that number, 
    - increases possibilities for error, 
    - and makes it more difficult to extend and adapt code in the future.  
  - Magic numbers should be replaced with constants or variables 
    - to make the program easier to maintain, understand, and read.
  + Use `students_per_group <- 4`
    - `num_groups <- 15`
    - `students <- students_per_group * num_groups`
  + Avoid `students <- 4 * 15`

Eliminate code duplication.  

  - Any block of code appearing in more than one m-file 
    - should be considered for packaging as a function.
  + It is much easier to manage changes and debug if code appears in only one file.  
  + Do not copy and paste blocks of code.  
    - If you find yourself cutting and pasting code, 
    - consider how you can create a new function or subfunction of this block of code.  

Make your code versatile.  

  + Consider how your code could be applied to different scenarios
  + Make sure there are error checks
  + Test your code (see below)
  * Use parentheses.  Parentheses help clarify operator precedence.  
  * Be careful of floating point comparisons

Here is an example.  

  - You can use the ```round``` or ```eps``` function in R 
    - to deal with these issues.
    
```{r}
# This is true
short_side <- 3
long_side <- 5
other_side <- 4
long_side ** 2 == (short_side ** 2 + other_side ** 2)

# But this is false
scale_factor <- 0.01
(scale_factor * long_side) ** 2 == ((scale_factor * short_side)
                                    ** 2 + (scale_factor * other_side) ** 2)

```

Write a test script for every function.  

  - Use the prefix ```test_``` for your test function 
    - and make use of the `assert` function.  
  - Write a test script for every function. 
    - This practice will improve the quality of the initial version  
    - and the reliability of changed versions. 
  - Consider that any function too difficult to test is probably too difficult to write. 

Boris Beizer: “More than the act of testing, 

  - the act of designing tests is one of the best bug preventers known.”
  + For example, to test the function fibonacci, 
    - you should write ```TestFibonacci()```

Functions should be kept short 

  - (guideline of under 30 lines of code).  

Modularize

  + Modularization enhances readability, understanding and testing 
    - by reducing the amount of text which must be read 
    - to see what the code is doing.  
  + Code longer than two editor screens is a candidate for partitioning. 
    - Small well designed functions are more likely 
    - to be usable in other applications.
  * Use existing functions.
  + It may be quicker or surer to find an existing function 
    - that provides some or all of the required functionality.

Vectorize

  + If you can program the same operations without the use of a ```for``` loop, 
    - then the code should run a lot faster.
  + Make use of ```lapply()```, ```sapply()```, and ```tapply()```
  - Or better yet, use the Tidyverse


##### Naming

- Consistency is very important in naming.  

  - The style of a name immediately indicates what the entity is:  
    - a function, variable, etc...

Use as descriptive a name as possible.  

  - Variables should be nouns, 
    - while functions should be "command" verbs. 

Do not try to save horizontal space, 

  - as it is far more important to make your code 
    - immediately understandable to a reader.
  + Use ```num_errors``` or ```num_completed_connections```
  + Avoid ```n``` or ```nerr``
  + Function names should be imperative (that is they should be commands).  
  + Use ```OpenFile()```

Names of dimensioned variables and constants 

  - should usually have a units suffix.
  + Adding a unit suffix helps to avoid confusion
  + Use ```incident_angle_radians``` not ```incident_angle```

NASA lost its \$125 million Mars Climate Orbiter 

  - because of issues with units in their data

Make sure there are no abbreviations in names

  + Use whole words to reduce ambiguity
  + Use ```ComputeArrivalTime()```
    + Avoid ```ComputeArr()```
  + Specific phrases that are naturally known through abbreviations 
    - should be kept abbreviated (e.g., cm).  
  - Even these cases might benefit from a defining comment 
    - at their first appearance. 

Names should be pronounceable 

  + Pronounceable names are easier to remember

##### Functions

* CamelCase is preferred for function names.

The prefix `Is` should be used for boolean functions

  + `IsOverpriced()`
  * There are a few alternatives to the `Is` prefix 
    - that fit better in some situations. 
  - These include the `Has`, `Can`, and `Should` prefixes:
    + `HasLicense()`, `CanEvaluate()`, `ShouldSort()`

Functions should have meaningful names

  + Make your filenames very specific.  
    - For example, use `http_sever_logs` rather than `logs`
  + Use `ComputeTotalWidth()` not `CompWid()`
  + An exception is the use of abbreviations or acronyms widely used.  
    - Functions with such abbreviations or acronyms 
    - should have the complete word in the first header line for clarity
  * Functions with a single output can be named for the output
    + Examples are `Mean()`, `StandardError()`
  * Functions with no output should be named for what they do
    + An example is `Plot()`

The prefix `Find` can be used in methods where something is looked up. 

  + Examples are `FindOldestRecord()`; `FindHeaviestElement()`
  * Complement names should be used for complement operations
    + Get/Set, 
    - Add/Remove/Create/Destroy, 
    - Start/Stop, 
    - Insert/Delete, 
    - Increment/Decrement, 
    - Old/New, 
    - Begin/End, 
    - First/Last, 
    - Up/Down, 
    - Min/Max, 
    - Next/Previous, 
    - Open/Close, 
    - Show/Hide, 
    - Suspend/Resume, 
    - etc...

Use explicit `return`

  - Use 
    - `AddValues <- function(x, y) {`
    - `return(x+y)`
    - `}`
  - Not 
    - `AddValues <- function(x, y) {`
    - `x+y`
    - `}`
  

##### Variable Names

* The names of variables should document their meaning or use.

The prefix `num` should be used 

  - for variables representing the number of objects.  
    + `num_files`, `num_segments`

The suffix `array` should be used 

  - for the pluralization of a variable
    + `point_array`

Variables representing a single entity number 

  - can be suffixed by Number.
    - `table_number`, `employeee_number`

Iterator variables should be named or prefixed with i, j, k etc.

  + `for (i_file in Mylist) {}`

Avoid negated Boolean operators 

  + Use `is_found` not `is_not_found`


#### Comments in your actual code blocks

- Comments are good and should be used.  

  - But they can sometimes be used to cover up for badly written code.  
  - Many times, heavily commented code is being used 
    - only because it is compensating for bad style.  
  - If you write your code properly, 
    - code comments are might actually superfluous.  
  - Before commenting your code, 
    - see if you can write your code in a way 
    - so that the comments aren't even needed.  

Example:

  - The following code uses magic numbers, 
    - and then explains what the magic numbers are through comments: 

```
solar_decom <- lapply(time_series, stlplus, s.window = "periodic", n.p = 24*4)
# The season for the data is one day because the cycle repeats over the course
# of 24 hours. Therefore, the frequency would be 24 hours *  4 measurements 
# per hour for a total of 96 measurements in a day.
```
Instead, the following code 

  - is more readable 
    - and reduces the need for comments:

```
hours_per_day <- 24
measurements_per_hour <- 4 # every 15 minutes
solar_decom <- lapply(time_series, stlplus, s.window = "periodic", n.p = hours_per_day*measurements_per_hour)
```



#### Comments

- In order for us to better understand our own work 

  - and to better to be able to share knowledge with your "customer"
  - is it important to comment your code! 

Even with the most stringent of commenting, 

  - we have returned to scripts with confusion and frustration.

As we like to say it: be kind to your future self!

Therefore, from now on **YOU MUST BE ACTIVELY COMMENTING YOUR CODE**. 

We will also be checking your adherence to the 80 character line limit. 

A comment is not helpful if it gets cut off halfway through explaining your work.

**NOTE: DO NOT WRITE THE TEXT-BASED ANSWER AS A COMMENT!**

#### Questions and Answering Style

- One word answers will now not receive full credit. 

  - But, you do not need to write an excessive amount for each question. 
  - Your comments and visuals should help us understand what you have done 
    - to solve the problem 
  - and the text portion is for you 
    - to explain the results / reasoning from the analysis. 

To make your life easier, 

  - we have included 'ANSWER ->' in places where 
    - we expect you to type your reasoning.

We want you to fully explain your reasoning 

  - and how the data/analysis leads you to these conclusions. 

If you feel like you need an additional 'ANSWER ->' 

  - feel free to add an additional space 
  - for your explanations! Be thorough!

This will enable us to give partial credit where possible.

Make sure to show the relative code outputs to the questions asked. 

Ask yourself if your submission can be read by anyone! 

**NOTE: DO NOT WRITE A PARAGRAPH FOR AN ANSWER.**

It would be easier if you have a text answer right below the question/sub-question.

If you decide to answer everything in one place, 

  - make sure you are referring to the correct sub-question.

We have observed that some students 

  - are not able to identify the sub-questions to be answers 
  - because they are lost in between the description of the question.

Please double check before you submit 

  - and use `Ctrl+ Shift+ O` 
  - to view all the important sections and sub-sections.

**NOTE**

The code blocks and comments in the assignment R Markdown files 

  - are merely a suggestion of what you can do. 

The commments are there to guide your process. 

However, there are many approaches to problems as mentioned before 

  - so your method may not always match and that's okay! 

#### Plots

- Data visualization is one of the key components of data science and 

  - it is critical that you use proper graphical representation of your analysis.

While this is a topic that will be dealt with in the next few weeks, 

  - it is a good idea to refer back to this section at a later stage.

While you can use base functions in R like plot(), 

  - for quick and dirty EDA plots for your personal use
  - we want you to be creative and 
    - customize professional-looking plots in your submissions.

You won't get full credit 

  - for the base output of the ggplot2:ggplot() function either. 

Experiment with different aspects of the ggplot2 package 

  - and make your plots more colorful and easy to follow.

The expectations of a good plot include: 

  - adding a plot title, 
  - axes labels (with units, if any), 
  - changing theme specifications, 
  - adding a meaningful axes scale 
  - and using good color schemes with appropriate legends.

An example of a professional plot is included below. 

**NOTE** 

We do not mean to make you think 

  - that the more complicated your plot is, 
    - the better. 

Some things we will be looking for are 

  - axis labels, 
  - units, 
  - titles, 
  - legends, etc.

Visuals should be purposeful. 

If your visuals are too complicated 

  - they can become distracting. 

Infact, you can start collecting useful code snippets 

  - as either, .R scripts 
  - or better yet, R Markdown .Rmd files. 

```{r}
library(ggplot2)

#using penguins dataset
peng_df <- palmerpenguins::penguins

mytheme <- theme(
  text = element_text(size = rel(3.5)),
  strip.text.y = element_text(size = rel(3.5)),
  strip.text.x = element_text(size = rel(3.5))
)
#sets a relative size for all labels
#Strip text is used when plots are faceted

mytheme2 <-
  theme(
    axis.text.x = element_text(
      size = 10,
      angle = 45,
      
      vjust = 0.5
    ),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14)
  )
#sets individual sizes
#ngle determines the angle of the text


ggplot(peng_df) + #calls df
  geom_violin(aes(
    x = as.factor(species),
    y = bill_length_mm,
    color = as.factor(species)
  )) +
  #creates vioplot
  geom_boxplot(aes(x = as.factor(species),
                   y = bill_length_mm),
               width = .05) +
  #Notice I did not add a color here
  #I think it looks bad
  #creates boxplot narrow
  labs(x = 'Species',
       y = 'Bill Length (mm)') + #x & y labs
  mytheme2 +
  #12pt 45degreex .5vjustx
  theme(legend.position = 'none')

## Versus
ggplot(peng_df) +
  geom_violin(aes(x = as.factor(species),
                  y = bill_length_mm)) +
  geom_boxplot(aes(x = as.factor(species),
                   y = bill_length_mm))

```

#### Grading

- With this information in mind, 

  - we would go through each and every code block 
  - and read through your comments as well as your answers.

Please be aware that you submit the right set of files 

  - as mentioned on Canvas submission page. 
  
We find it difficult to follow up individually 

  - about individual submissions 
    - and tracking re-submissions. 
  - Please avoid such situations.

We would be following an absolute grading system 

  - in which we grade solely on the answers 
  - of what a person has included in their submission.

Your score is reflective of your performance on the lab exercise.

Please be mindful that the grading is done on a curve, 

  - given the diversity of this course.

That way, an A does not mean 90+ points on the course!

#### Shortcuts in RStudio

- To view the document outline 

  - and what all questions to answer (as mentioned earlier): 
    - `Ctrl + Shift + O`

To reformat code: 

  - `Ctrl + Shift + A`

Please make sure you use these shortcuts each time you work on the code.

#### Problems with Virtual Computing

- As mentioned earlier, 
  - please send a ticket explaining your problem 
    - and the class details 
    - (that you are a student of DSCI353/353M/453) to:-
    - help@case.edu: if you are facing issues with the ODS VDI

hpcsupport@case.edu: 

  - if you are facing issues with Markov/RStudio Server 

#### Final Thoughts

- The best way to learn this course 

  - is not only by learning on your own 
    - but also by talking to your peers 
    - and communicating with the professor as well as the TAs.

While discussing ideas with your peers is highly encouraged, 

  - plagiarism will not be tolerated.

We want you to make the best use of this course 

  - by working sincerely 
  - and being open to learning new concepts.

As data science concepts can be applied to any industry you can think of, 

  - please make sure you are paying attention to the coursework 
  - and learning as much as you can.

Please contact us if you face any difficulty in the course 

  - and we'd be more than happy to help you out 
    - (this encompasses a range of problems 
    - from course difficulty to RStudio issues).

Even if you think your questions are silly and not worth asking, please ask.

  - We'll try our best to answer them!

We will be on Slack most of the times and have regular office hours. 

We'll try and be around on weekends if you have questions.

Before you go, here's the last section you'd want to pay attention to 

  - (caution: there will be some scrolling 
    - but you'll see why it can be annoying 
    - to have the entire dataset printed).

#### Code Chunks

- We will first be more closely looking at the outputs of code chunks. 

Let's take a look at the following.

```{r}
library(tidyverse)
library(palmerpenguins)

df <- palmerpenguins::penguins

as.data.frame(df)

```

Here, we have forced the knitted file to express the entire dataset, a full 79 pages! 

Additionally when tidyverse was used in R 

  - it expressed a bunch of stuff 
    - we don't want in our final report! 

We do not want to see either of these examples in a submitted assignment anymore! 

If you need to print a data frame or value 

  - make sure that it does not take up pages. 

We suggest you use dplyr::glimpse() or dplyr::tibble() 

  - which both natively suppress base::print() statements 
    - to just a few rows. 

Notice that we are declaring the namespace when we code. 

  - There are thousands of packages, 
  - so it is often necessary what package's function you are using! 

We can suppress WARNINGS and MESSAGES in our finished report 

  - by adding it to the beginning of our code blocks like so:- 

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(palmerpenguins)
df <- palmerpenguins::penguins

df %>%
  dplyr::group_by(year) %>%
  dplyr::tally()

##Print Statements should go at the end of a code block
dplyr::tibble(df)
dplyr::glimpse(df)


```

That's it! Hope this gives you a better idea of our expectations early on.

Happy learning and hope you all do well!


