---
title: "CWRU DSCI351-351M-451: Week04a(CWRU, Pitt, UCF, UTRGV)"
subtitle: "Profs: R. H. French, L. S. Bruckman, P. Leu, K. Davis, S. Cirlos"
author: "TAs: W. Oltjen, K. Hernandez, M. Li, M. Li, D. Colvin" 
date:  "`r format(Sys.time(), '%d %B, %Y')`"  
output: 
  pdf_document:
    latex_engine: xelatex
    toc: true
    number_sections: true
    toc_depth: 6
    highlight: tango
  html_document:
    toc: yes
urlcolor: blue
---

 \setcounter{section}{4}
 \setcounter{subsection}{1}
 \setcounter{subsubsection}{2}

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  cache = FALSE, # if TRUE knitr will cache results to reuse in future knits
  fig.width = 6, # the width for plots created by code chunk
  fig.height = 4, # the height for plots created by code chunk
  fig.align = 'center', # how to align graphics. 'left', 'right', 'center'
  dpi = 300, 
  dev = 'png', # Makes each fig a png, and avoids plotting every data point
  # eval = FALSE, # if FALSE, then the R code chunks are not evaluated
  # results = 'asis', # knitr passes through results without reformatting
  echo = TRUE, # if FALSE knitr won't display code in chunk above it's results
  message = TRUE, # if FALSE knitr won't display messages generated by code
  strip.white = TRUE, # if FALSE knitr won't remove white spaces at beg or end of code chunk
  warning = FALSE, # if FALSE knitr won't display warning messages in the doc
  error = TRUE) # report errors
  # options(tinytex.verbose = TRUE)
```


#### Some simple settings suggestions for windows or mac

  - In Windows File Manager, make sure (show file extensions) is checked
  - On a Mac, no Insert Key, so Shift Insert doesn't work
    - Use two-finger tap on trackpad, to get popup menu
    - Then choose paste

#### An example of EDA with pipes

- Library in the packages we will use

```{r}
library(ggplot2) # load package

? ggplot2

library(dplyr)  #load package

? dplyr
```


##### Diamonds dataset

- Do load the Diamonds dataset and do some quick EDA

```{r}
data(diamonds)  # load dataset that comes with ggplot

head(diamonds) # Return the First or Last Part of an Object
```

```{r}
str(diamonds) # Compactly Display the Structure of an Arbitrary R Object
```

```{r}
glimpse(diamonds)  # the tidyverse version of str
```


```{r}
summary(diamonds) #  produce result summaries of the results of
# various model fitting functions
```

##### Now some compact ggplot2 EDA code

- Use the pipe `%>%` operator  

  - Which simply passes the output of the left operator 
  - As the first argument to the right operator

```{r}
diamonds %>%
  ggplot(aes(x = carat, y = price)) +  # aes is aesthetic mapping
  geom_point(alpha = 0.5) + # each data point as a point
  facet_grid( ~ cut) + # facet the scatter plot on cut, color or clarity
  stat_smooth(method = lm, formula = y ~ poly(x, 2)) + # fit a 2nd order lin. model
  theme_bw()
```

With this simple visualization, 

  - We can quickly see that price increases with carat size,
    - The relationship is nonlinear, 
  - There are some outliers, 
    - And the relationship does not depend too heavily on cut. 

###### Now lets use GGally and GGpairs packages

- These packages are extensions to GGplot 2

```{r}
library(GGally) # a ggplot2 extention; gallery of plot templates

? GGally  # This doesn't work for the GGally package

?? GGally # So try this
```

```{r}
diamonds %>%
  mutate(volume = x * y * z) %>%    # in the pipe calculate the volume
  select(cut, carat, price, volume) %>%
  sample_frac(0.5, replace = TRUE) %>%
  ggpairs(axisLabels = "none") +
  theme_bw() 
```

The Tidyverse functions 

  - mutate, select, and sample_frac (verbs!) 
  - Are part of the dplyr data manipulation library
  - And are loaded when you load the tidyverse metapackage

#### References

  - Deep Ganguli [The Grammar of DSCI](http://technology.stitchfix.com/blog/2015/03/17/grammar-of-data-science/)

