---
title: "DSCI351-351m-451: LE "
subtitle: "2200-353-353m-453-LE2a"
author: "R. H. French, L. S. Bruckman, P. Leu, K. Davis, S. Cirlos"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    latex_engine: xelatex
    toc: true
    number_sections: true
    toc_depth: 6
    highlight: tango
  html_document:
    toc: yes
urlcolor: blue
---

<!--
# Script Name: 
# Authors: Roger H. French
# License: Creative Commons Attribution-ShareAlike 4.0 International License.
##########
# Latest Changelog Entires:
# v0.00.01 - 1501cwru-dsci-NAMEIT.RMD - Roger French started this blank Rmd
##########

# Rmd code goes below the comment marker!
-->


\setcounter{section}{2}
\setcounter{subsection}{1}
\setcounter{subsubsection}{0}

#### Lab Exercise 2a: Classifiation

2 Questions, for a total of 2 points

  - Question 1 = 1.5 points
  - Question 2 = 0.5 points

#### Question ISLR Chapter 4, Exercise 10


    
##### (a) 

```{r}

```

Answer: 



#### Question ISLR Chapter 4, Exercise 12



##### (a) Write a function, Power(), 

  
  
```{r}

```

Answer:   



#### Links
 


<!--
# Keep a complete change log history at bottom of file.
# Complete Change Log History
# v0.00.00 - 1405-07 - Nick Wheeler made the blank script
##########

