---
title: "w03b p2 Coding Expectations and Clean R Code"
ssubtitle: "Profs: R. H. French, L. S. Bruckman, P. Leu, K. Davis, S. Cirlos"
author: "TAs: W. Oltjen, K. Hernandez, M. Li, M. Li, D. Colvin" 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    latex_engine: xelatex
    toc: TRUE
    number_sections: TRUE
    toc_depth: 6
    highlight: tango
  html_notebook:
  html_document:
    css: ../lab.css
    highlight: pygments
    theme: cerulean
    toc: yes
    toc_depth: 6
    toc_float: yes
    df_print: paged
urlcolor: blue
always_allow_html: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  cache = FALSE, # if TRUE knitr will cache results to reuse in future knits
  fig.width = 5, # the width for plots created by code chunk
  fig.height = 3.5, # the height for plots created by code chunk
  fig.align = 'center', # how to align graphics. 'left', 'right', 'center'
  dpi = 300, 
  dev = 'png', # Makes each fig a png, and avoids plotting every data point
  # eval = FALSE, # if FALSE, then the R code chunks are not evaluated
  # results = 'asis', # knitr passes through results without reformatting
  echo = TRUE, # if FALSE knitr won't display code in chunk above it's results
  message = TRUE, # if FALSE knitr won't display messages generated by code
  strip.white = TRUE, # if FALSE knitr won't remove white spaces at beg or end of code chunk
  warning = FALSE, # if FALSE knitr won't display warning messages in the doc
  error = TRUE) # report errors
  # options(tinytex.verbose = TRUE)
```

\setcounter{section}{3}
\setcounter{subsection}{2}
\setcounter{subsubsection}{3}

#### Coding Expectations in DSCI courses

- Congratulations! You're working on lab exercise 1 (LE1). 

Shortly you'll receive your LE scores and feedback about their submissions. 

This document outlines general expectations moving forth.

Some of you who took the DSCI course last semester 

  - may have seen a similar document.

Please read through this document carefully 

  - and if you have any questions,
do message us on Slack.

##### General Expectations

- Please follow these points carefully and work on your lab exercises.

The LE grading will be done based on these points

To reiterate the points:-

  - Remember that the knitted PDF/HTML shows the first 80 characters in the code
  as the output. 
  - And using Canvas grading, we grade on your pdf
    - Make sure you format your code chunks properly. 
    - Please verify this before submission.
  
  - Since we'll be grading your work, be sure to comment your code. 
    - Suppose you use a function that is not common (for example, unite()), 
    - explain what it does in a short comment 
    - and which package you used. 
  - Best way is to define a function and package as follows
    -  `ggplot2::ggsave`
    - This defines the package and function you are using
  - It would also help me learn about new functions in the process. 
  
  - There are multiple approaches to the same question. 
    - we'll take that into account while grading.
  
  - Please answer directly under subquestions (indicated by -). 
  - If you want to answer everything in one place, 
    - say which subquestion you are referring to. 
    - For example: Answer to 'what's the highest ranked...': answer here.
  
  - Please answer questions outside the code chunk. 
    - Suppose, the question asks you to compare trends, 
    - don't type it as a code comment.
  
  - Be creative while you work on plots. 
    - ggplot2 package has amazing functionalities. 
    - Spend some time formatting plots to make them more understandable. 
  - For example, 
    - adding a plot title, 
    - axes labels (specify units wherever required), 
    - changing theme specifications will earn full points.
  
  - Include meaningful variable names. 
  - e.g. for the billboard lyrics dataset, 
    - billboard_lyrics is too long and bbl is too short. 
    - Find a balance between the two extremes like 'songs'. 
  
Your see our comments in your lab exercises when graded. 

  - If there are any questions on these aspects, 
  - don't hesitate to contact us on Slack.

##### Slack

- From now on, please use Slack as your primary mode of communication with us, 

  - especially for asking questions. 

Don't forget to tag us by using @name 

  - because it's highly probable that your messages might get lost before the submission. 

While asking questions, refer to the question number in your question.

Beaware of threads in the slack channel.

Ask new questions, outside prior hreads 

  - so that we won't miss them.

Ask your questions in the class channel #dsci353-353m-453. 

  - This way others can also benefit and help with your answers. 
  - It'll be a win-win situation!

If something has to be privately discussed with us or Prof. French, 

  - please contact by DM, on 1on1 message in slack

##### Office Hours

- We have two office hours in the week:-

  - Mondays and Wednesdays from 4-5 pm.

Please attend the office hours if you have questions and any other concerns.

Generally these sessions would be recorded.

##### Submission Time

- Typically lab exercises are expected to be submitted by 11:59 pm of the deadline day.

If you are facing any challenges or problems 

  - that prevent you from turning in the lab exercise on time, 
  - please contact us with an explanation of why 
    - we'll see if your request can be accommodated.

##### Git Commands

- **NOTE: NOT FOLLOWING THE GIT COMMANDS COULD LEAD TO SEVERAL PROBLEMS.**

I cannot emphasize enough why this is important.

Personal story (some of you might know about this): when I was a DSCI student, I 
wasn't regular at git pushing and git pulling.

At one point, I lost all of my files from the ODS VDI and thankfully, it was 
possible to retrieve after spending a few hours from the CWRU Network.

All of this happened just before the midterm exam!

It was definitely scary.

Please don't land in such situations!

Git pushing and pulling regularly will save your work virtually and in events of
your computer crashing with no particular reason, you can always clone your fork.

If you are not doing it already, please make it a habit.

##### Markov Data Science Cluster, vs ODS Win10 Desktop Versus Local Computer

- There are several reasons why we encourage you to work on the ODS VDI for a couple
of lab exercises. 

This is because the packages would be updated from time to time and it's easier
to reach out to help@case.edu about the problems. 

In fact, we can try diagnosing the issues from our Markov too.

In such cases, make sure you have the Git set up 

  - and do git pushing/pulling regularly.

If you ever face any issue, please contact us.


##### Coding Style

- Code styling is important!

  - As data scientists, part of our job is effectively communicating our process
  and results to a broader audience. 
  
We expect your code to be 

  - compact, 
  - well-commented and 
  - effectively answers the question you are trying to solve.

There needs to be a whitespace between symbols and characters. 

  - `hits %>% filter(Artist=="madonna")` is bad style
  - `hits %>% filter(Artist == "madonna")` is good style

There are multiple methods to solve a question 

  - and new approaches to solving a question would always be appreciated.
  
If you are using a new function that is not regularly used, 

  - please use package::function approach and say what it does. 

Thus, going forward we will 

  - scrutinize the following code style things 
  - and points will be deducted accordingly. 

**NOTE**

When looking at the `.Rmd` file you can notice something interesting. 

Every sentence I write is ended by a **PERIOD** a **SPACE** and a **ENTER** key.

Although it creates a new line of code, 

  - when it is knitted they will all be in one block of text. 
  
To make a separate paragraph

  - requires a blank line

##### Code Chunks

- We will first be more closely looking at the outputs of code chunks. 

Let's take a look at the following.

```{r}
library(tidyverse)
library(palmerpenguins)

df<-palmerpenguins::penguins

as.data.frame(df)

```

Here, I have forced the knitted file to express the entire dataset, a full 79 pages! 

Additionally when tidyverse was libraried into R 

  - it expressed a bunch of stuff we may not want in our final report! 

We do not want to see either of these examples 

  - in a submitted assignment anymore! 

If you need to print a data frame or value make sure that it doesnt take up pages. 

We suggest you use 

  - `dplyr::glimpse()` 
  - or `dplyr::tibble()` 
  - which both natively suppress `base::print()` statements to just a few rows. 

Notice that **I am declaring the namespace** in my code, 

  - there a thousands of packages 
  - and people aren't that creative 
    - so it is often necessary what package's function you are using! 

We can suppress WARNINGS and MESSAGES in our finished report 

  - by adding it to the beginning of our code blocks like so: 

  - ````{r, warning = FALSE, message = FALSE}` 

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(palmerpenguins)
df <- palmerpenguins::penguins

df %>% 
  dplyr::group_by(year) %>% 
  dplyr::tally()

##Print Statements should go at the end of a code block
dplyr::tibble(df)
dplyr::glimpse(df)


```

##### Comments

- In order for us to better understand our own work and to better to be able to 
share knowledge is it important to comment your code! 

Even with the most strigent of commenting, 

  - I myself have returned to scripts written with confusion and fustration. 

Therefore, from now on **YOU MUST BE ACTIVELY COMMENTING YOUR CODE**. 

We will also be checking your adherence to the 80 character line limit. 

A comment is not helpful if it gets cut off halfway through explaining your work.

**NOTE: DO NOT WRITE THE TEXT-BASED ANSWER AS A COMMENT!**

##### Questions and Answering Style

- One word answers will now not receive full credit. 

We want you to fully explain your reasoning 

  - and how the data/analysis leads you to these conclusions. 

This will enable us to give partial credit where possible. 

Make sure to show the relative code outputs to the questions asked. 

Ask yourself if your report can be read by anyone! 

**NOTE: DO NOT WRITE A PARAGRAPH FOR AN ANSWER.*

It would be easier if you have a text answer 

  - right below the question/sub-question.

If you decide to answer everything in one place, 

  - make sure you are referring to the correct sub-question.

Please double check before you submit.

**NOTE**

The code blocks and comments in the assigment R Markdown files 

  - are merely a suggestion, or guidance, of what you can do. 

The comments are there to guide your process. 

However, there are many approaches to problems 

  - as mentioned before so your method
  - may not always match and that's okay! 

##### Plots

- Plots now will also have stricter requirements. 

You will no longer receive full credit for the base output 

  - of the ggplot2:ggplot() function. 
  - or for using the base `plot` command

Experiment with different aspects of the ggplot2 package 

  - and make your plots more colorful and easy to follow.

The expectations of a good plot include: 

  - adding a plot title, 
  - axes labels (with units, if any), 
  - changing theme specifications, 
  - adding a meaningful axes scale 
  - and using good color schemes with appropriate legends.

An example of a professional plot is included below. 

**NOTE** 

We do not mean to make you think 

  - that the more complicated your plot is, 
  - the better. 

Visuals should be purposeful. 

If your visuals are too complicated 

  - it can become distracting. 

With that being said, 

  - you are highly encouraged to reuse past work for future submissions. 

Infact, you can start collecting useful code snippets in 

  - either, .R scripts or better yet, R Markdown files. 

```{r}
library(ggplot2)

df <- palmerpenguins::penguins

mytheme <- theme(
  text = element_text(size = rel(3.5)),
  strip.text.y = element_text(size = rel(3.5)),
  strip.text.x = element_text(size = rel(3.5))
)
#sets a relative size for all labels
#Strip text is used when plots are faceted

mytheme2 <-
  theme(
    axis.text.x = element_text(
      size = 10,
      angle = 45,
      
      vjust = 0.5
    ),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14)
  )
#sets individual sizes
#Angle determines the angle of the text


ggplot(df) + #calls df
  geom_violin(aes(
    x = as.factor(species),
    y = bill_length_mm,
    color = as.factor(species)
  )) +
  #creates vioplot
  geom_boxplot(aes(x = as.factor(species),
                   y = bill_length_mm),
               width = .05) +
  #Notice I did not add a color here
  #I think it looks bad
  #creates boxplot narrow
  labs(x = 'Species',
       y = 'Bill Length (mm)') + #x & y labs
  mytheme2 +
  #12pt 45degreex .5vjustx
  theme(legend.position = 'none')

## Versus
##
##
ggplot(df) +
  geom_violin(aes(x = as.factor(species),
                  y = bill_length_mm)) +
  geom_boxplot(aes(x = as.factor(species),
                   y = bill_length_mm))

```

##### Grading

- With this information in mind, 

  - We would go through each and every code block 
  - and read through your comments as well as your answers.

Your score is reflective of your performance on the lab exercise.

Please be mindful that the grading is done on a curve, 

  - given the diversity of the students in the course.

That way, an A does not mean 90+ points on the course!

##### Shortcuts in RStudio

- To view the document outline and what all questions to answer: Ctrl + Shift + O

To reformat code: Ctrl + Shift + A

Please make sure you use these shortcuts each time you work on the code.

##### Final Thoughts

- The best way to learn this course is not only by learning on your own 

  - but also by talking to your peers 
  - and communicating with the professor as well as the TA.

While discussing ideas with your peers is highly encouraged, 

  - plagiarism will not be tolerated.

We want you to make the best use of this course 

  - by working sincerely 
  - and being open to learning new concepts.

As data modeling concepts can be applied to any industry you can think of, 

  - please make sure you are paying attention to the coursework 
  - and learning as much as you can.

Please contact us if you face any difficulty in the course 

  - and we'd be more than happy to help you out 
  - (this encompasses a range of problems from course difficulty to RStudio issues).

We will be on Slack most of the times 

  - and have regular office hours. 

--------------------------------------------------------


Happy learning and hope you all do well!

#### Clean R Code Is Critical \

- Over many years of experience delivering successful projects, 

  - There is one common element across all these projects 
  
A clean, readable, and concise code base 

  - is the key to effective collaboration 
  - and provides the highest quality value to the client.

##### Code Review \

- Code review is a crucial part 

  - of maintaining a high-quality code process. 
  - It is also a great way to 
    - share best practices 
    - and distribute knowledge among team members. 
  - Code review as a must for every project. 
    - Lets review best practices 
    - recommended for all data science teams.

Having a well-established code review process does not change the fact 

  - that the data scientist is responsible for
    - writing good, clean code! 
  - Pointing out all of the code’s basic mistakes 
    - is painful, time-consuming, 
  - And distracts reviewers from going deep 
    - into code logic 
    - or improving the code’s effectiveness.

Poorly written code can also harm team morale 

  - code reviewers are frustrated 
    - while code creators might feel offended by a huge number of comments. 

That is why before sending the code to review, 

  - developers need to make sure that the code is as "clean" as possible. 

Also, note that there is not always a code reviewer that can come to the rescue. 

  - Sometimes you are on your own in a project. 
  - Even though you think the code is ok for you now, 
    - consider rereading it in a few months 
    - you want it to be clear to avoid wasting your own time later on.

Lets summarize 

  - the most common mistakes to avoid 
  - and outline best practices to follow 
    - in programming in general. 
  - Follow these tips to speed up the code review iteration process 
    - and be a better data scientist

##### Avoid Comments with Comments \

- Adding comments to the code is a crucial developer skill. 

  - However, a more critical and harder to master skill 
    - is knowing when not to add comments. 
  - Writing good comments is more of an art than a science. 
  - It requires a lot of experience, 
    - and you can write entire book chapters about it 
    - (e.g., [Clean Code: A Handbook of Agile Software Craftsmanship](https://books.google.pl/books/about/Clean_Code.html?id=hjEFCAAAQBAJ). 

There are few simple rules that you should follow, 

  - to avoid comments about your comments:

The comments should add external knowledge to the reader: 

  - if they’re explaining what is happening in the code itself, 
    - it is a red flag that the code is not clean 
    - and needs to be refactored. 

##### Code Refactoring \

- What is **[code refactoring](https://en.wikipedia.org/wiki/Code_refactoring)**

  - code refactoring is the process of restructuring existing computer code
    - changing the factoring
    - without changing its external behavior. 
  - Refactoring is intended to improve 
    - the design, 
    - structure, 
    - and/or implementation of the software 
    - (its non-functional attributes), 
  - while preserving its functionality. 

Potential advantages of refactoring may include 

  - improved code readability 
    - and reduced complexity; 
  - these can improve the source code's 
    - maintainability 
  - and create a 
    - simpler, 
    - cleaner, 
    - or more expressive internal architecture 
    - or object model to improve extensibility. 
  - Another potential goal for refactoring is improved performance; 
    - software engineers face an ongoing challenge 
    - to write programs that perform faster 
    - or use less memory.

##### Comments 

- So in your code comments, if some hack was used, 

  - then comments might be used to explain what is going on. 
  - Comment required business logic 
    - or exceptions added on purpose. 
  - Try to think of what can be surprising to the future reader 
    - and preempt their confusion.
  - Write only crucial comments! 
    - Your comments should not be a dictionary of easily searchable information. 
    
In general, comments are distracting 

  - and do not explain logic as well as the code does. 

For example, I recently saw a comment like this in the code: 

  - `trimws(.) # this function trims leading/trailing white spaces` 
    - This comment is is redundant. 
  - If the reader does not know what function trimws is doing, 
    - it can be easily checked. 
  - A more robust comment here can be helpful, 
    - e.g.: `trimws(.) # TODO(Marcin Dubel): Trimming white spaces is crucial here due to database entries inconsistency; data needs to be cleaned.`

##### Use `roxygen2` inline documentation 

- When writing functions in R, I recommend using {`roxygen2`} comments 

  - even if you are not writing a package. 
  
```{r}
library(roxygen2)
?roxygen2
```

`roxygen2` is a package used for building R packages

  - Generate your 
    - Rd documentation, 
    - 'NAMESPACE' file, 
    - and collation field 
  - using specially formatted comments. 
  - Writing documentation in-line with code 
    - makes it easier to keep your documentation up-to-date 
    - as your requirements change. 
  - 'Roxygen2' is inspired by the 'Doxygen' system for C++.
  - Python3 has `[sphinx](https://en.wikipedia.org/wiki/Sphinx_(documentation_generator))` 


`roxygen2` is an excellent tool for organizing the knowledge about 

  - the function goal, 
    - parameters, 
    - and output.

##### More on commenting 

- Only write comments (as well as all parts of code) in English. 

  - Making it understandable to all readers 
    - might save you encoding issues that can appear 
    - if you use special characters from your native language.

In case some code needs to be refactored/modified in the future, 

  - mark it with the # TODO comment. 
  
Also, add some information 

  - to identify you as the author of this comment 
    - (to contact in case details are needed) 
  - and a brief explanation of 
    - why the following code is marked as TODO 
    - and not modified right away.

Never leave commented-out code un-commented! 

  - It is ok to keep some parts for the future 
    - or turn them off for a while, 
    - but always mark the reason for this action.
    
Remember that the comments will stay in the code. 

  - If there is something that you would like to tell your reviewer, 
    - but only once, 
  - add a comment to the Pull (Merge) Request 
    - and not to the code itself.

Example: I recently saw removing part of the code with a comment like: 

  - “Removed as the logic changed.” 
  - Ok, good to know, 
    - but later that comment in the code looks odd and is redundant, 
    - as the reader no longer sees the removed code.


##### Strings 

- A common problem related to texts 

  - is the readability of string concatenations. 
  - What one encounters a lot 
    - is an overuse of the paste function. 
  - Don’t get me wrong; 
    - it is a great function when your string is simple, 
    - e.g. `paste("My name is", my_name)`, 
  - but for more complicated forms, it is hard to read: 

```{}
paste("My name is", my_name, "and I live in", my_city, "developing in", 
      language, "for over", years_of_coding)
```
A better solution is to use 

  - `sprintf` functions 
  - or `glue`, e.g. 

```{}
glue(“My name is {my_name} and I live in {my_city} developing in {language}
     for over {years_of_coding}”)
```

Isn’t it clearer 

  - without all those commas 
  - and quotation marks?

When dealing with many code blocks, 

  - it would be great to extract them to separate locations, 
    - e.g., to a .yml file. 
  - It makes both code and text blocks 
    - easier to read and maintain.

The last tip related to texts: 

  - one of the debugging techniques, 
    - often used in Shiny applications, 
    - is adding `print()` statements. 
  - Double-check whether the prints are not left in the code 
    - this can be quite embarrassing during code review!

##### Loops 

- Loops are 

  - one of the programming building blocks 
  - and are a very powerful tool. 

Nevertheless, they can be computationally heavy 

  - and thus need to be used carefully. 

The rule of thumb that you should follow is: 

  - always double-check if looping is a good option. 

It is hardly a case that 

  - you need to loop over rows in data.frame: 
    - there should be a {`dplyr`} function 
    - to deal with the problem more efficiently. 

Another common source of issues is 

  - looping over elements 
    - using the length of the object, 
    - e.g. `for(i in 1:length(x)) ....` 
  But what if the length of x is zero! 
    - Yes, the loop will go another way 
    - for iterator values 1, 0. 
  - That is probably not your plan. 
    - Using `seq_along` or `seq_len` functions 
    - are much safer.

Also, remember about the `apply` family of functions for looping. 

  - They are great 
    - (not to mention {the `purrr` package} solutions)! 
  - Note that using `sapply` 
    - might be commented by the reviewer as not stable 
    - because this function chooses the type of the output itself! 
  - So sometimes it will be 
    - a list, 
    - sometimes a vector. 
  - Using vapply is safer, 
    - as the programmer defines the expected output class.

##### Code Sharing 

- Even if you are working alone, 

  - you probably would like your program 
    - to run correctly on other machines. 
  - And how crucial it is 
    - when you are sharing the code with the team! 
  - To achieve this, 
    - **never use absolute paths in your code**, 
    - e.g. `/home/marcin/my_files/old_projects/september/project_name/file.txt`. 
  - It won’t be accessible for others. 
    - Note that any violation of folder structure will crash the code. 

As you should already **have an Rproject for all coding work**, 

  - you need to use paths related to the particular Rproject
    - in this case; it will be `./file.txt`. 
  - What is more, **one would suggest keeping all the paths** 
    - as variables in a single place
  - so that renaming a file requires one change in code, 
    - not, e.g., twenty in six different files.

Sometimes your software needs to use some credentials or tokens, 

  - e.g., to a database or private repositories. 
    - or an external API, like Google Maps
  - You should never commit such secrets to the git repository! 
    - Even if the entries are the same among the team. 
  - Usually, the good practice is to keep such values 
    - in .Renviron file as environmental variables 
    - that are loaded on start 
    - and the file itself is ignored in the repo. 
    - You can [read more about .Renviron here](http://www.dartistics.com/renviron.html).
  - Or use the `keyring` package
    - It stores tokens or credentials
    - And exists for both R and Python3

##### Good Programming Practices 

- Finally, let’s focus on how you can improve your code. 

  - First of all, 
    - your code should be easily understandable and clean 
  - even if you are working alone, 
    - when you come back to code after a while, 
    - it will make your life easier! 

Use specific variable names, 

  - even if they seem to be lengthy 
  - the rule of thumb is that you should be able to guess 
    - what is inside just by reading the name, 
    - so `table_cases_per_country` is ok, 
    - but `tbl1` is not. 
  - Avoid abbreviations. 
    - Lengthy is preferable to vague. 
  - Keep consistent style for object names 
    - (like camelCase or snake_case) 
    - as agreed among your team members. 

Do NOT abbreviate logical values 

  - such as `T` for `TRUE` 
    - and `F` for `FALSE` 
  - the code will work, 
    - but `T` and `F` are regular objects 
    - that can be overwritten 
  - while `TRUE` and `FALSE` are special values
    -  as defined in R. 

Do not compare logical values using equations, 

  - like `if(my_logical == TRUE)`. 
  - If you can compare to `TRUE`, 
    - it means your value is already logical, 
    - so `if(my_logical)` is enough! 
  - If you want to double-check 
    - that the value is `TRUE` indeed 
    - (and not, e.g., NA), 
    - you can use the isTRUE() function.

Make sure that your logic statements are correct. 

  - Check if you understand the difference in R 
    - between single and double logical operators!

Good spacing is crucial for readability. 

  - Make sure that the rules are the same 
    - and agreed upon in the team. 
  - It will make it easier to follow each other’s code. 
  - The simplest solution is to stand on the shoulders of giants 
    - and follow the tidyverse style guide.
    - Its the same as the Google R style guide.

However, checking the style in every line 

  - during the review is quite inefficient, 
  - so make sure to introduce `lintr` and `styler` 
    - in your development workflow
  - Our use the code diagnostics in Rstudio 
  - This can be lifesaving! 
  
Recently we found an error in some legacy code 

  - that would have been automatically recognized by `lintr`:

```{}
sum_of_values <- first_element 
  + second_element
```

This does not return the sum of the elements 

  - as the author was expecting.

Speaking of variable names 

  - this is known to be one of the hardest things in programming. 
  - Thus avoid it when it is unnecessary. 

Note that R functions return, by default, 

  - the last created element, 
  - so you can easily replace that:

```{}
sum_elements <- function(first, second) {
  my_redundant_variable_name <- sum(first, second)
  return(my_redundant_variable_name)
}
```

With something shorter 

  - (and simpler, 
    - you don’t need to think about names):

```{}
sum_elements <- function(first, second) {
  sum(first, second)
}
```

On the other hand, please DO use additional variables 

  - anytime you repeat some function call or calculation! 
  - It will make it computationally more effective 
    - and easier to be modified in the future. 

Remember to keep your code DRY 

  - don’t repeat yourself. 
  - If you copy-paste some code, 
  - think twice whether it 
    - shouldn’t be saved to a variable, 
    - done in a loop, 
    - or moved to a function. 

##### Conclusion \

- And there you have it 

  - five strategies to write clean R code 
    - and leave your code reviewer commentless. 
  - These five alone will ensure you’re writing great-quality code 
    - that is easy to understand, 
    - even years down the road.



#### Links

  - Marcel Dubel, [Clean Code](https://appsilon.com/write-clean-r-code/)
  -  [Clean Code: A Handbook of Agile Software Craftsmanship](https://books.google.pl/books/about/Clean_Code.html?id=hjEFCAAAQBAJ)